---
title: "Complete"
author: "Amieroh Abrahams"
date: "11 August 2018"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 6.5
    fig_width: 6.5
    highlight: default
  html_document:
    fig_caption: yes
    fig_height: 7
    fig_retina: 2
    fig_width: 7
    highlight: espresso
    theme: cerulean
    toc_float: no
  md_document:
    variant: markdown_github
  word_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    highlight: zenburn
toc: yes
mainfont: PT Serif
monofont: PT Mono
language: Australian
sansfont: PT Sans
fontsize: 12pt
---
\newpage
## Plagiarism Declaration

\newpage
## Declaration

\newpage
## Abstract

The South African coastline is comprised of distinct coastal regions, each varying in temperature. Seawater temperature is an important driver affecting marine biodiversity, and so understanding the factors affecting the variations of this is important. Here, I analysed temperature time series data for 18 different sites distributed along the South African coastline. These data were collected using underwater temperature recorders (UTRs), alcohol thermometers, and satellites, with the precision of the instruments varying from between 0.5 ºC to 0.001ºC. Using the minimum, maximum and mean temperatures I systematically grouped sites together in their distinct coastal regions and compared temperatures between regions. I further assessed the effects of wind and wave exposure on the variations of temperatures using coefficient of determination analyses. My results showed that wind and wave action per site were not significantly affecting temperature, achieving low adjusted R2 values, indicating that wave and wind action were not the predominant factors influencing coastal seawater temperature. Large scale oceanic processes such as coastal upwelling and the presence of major ocean currents, as well as anthropogenically induced factors were more likely to be the main drivers affecting temperature. This suggests that human interreference may be indirectly influencing marine biodiversity via affecting ocean temperatures, and this insight could prove useful in aiding in conservation.

*Keywords*: Seawater temperature, climate change, coastal regions, code: R, variability

\newpage
## Acknowledgments

Special thanks go to my supervisor, AJ Smit, for the continuous motivation and guidance throughout my honours year. All of your contributions and efforts towards my academic career and experiences have been appreciated beyond measure. I would also like to thank Dr. Robert Schlegel and my co-supervisor, Dr. Robert Williamson for the help with my R analyses. I would like to thank KZNSB, DEA, DAFF, SAEON, SAWS, UWC and EKZNW for contributing the raw data allowing me to do this study. Without the data this thesis would not be possible.I would also like to thank my colleagues within Team Kelp who assisted me throughout this project. Their discussions and friendship has largely aided in the completion of this thesis. To my friends and family, thank you for providing me with the necessary support and encouragement. I would also like to acknowledge the entire Biodiversity and Conservation biology department at the University of the Western Cape. Lastly, i would like to thank the NRF for providing the necessary funding towards the completion of this project.

\newpage
# Introduction

Temperature is the main determinant of biogeographical patterns and ecosystem processes, having observable effects on an organism’s survival and reproductive patterns (Smit et al. 2013; Pearce et al. 2006; Smale and Wernberg, 2009). The temperatures of seawater are a key indicator of environmental changes in marine ecosystems (Pearce et al. 2006). Despite the controlling effects that temperature plays on the growth, survival, and distribution of marine organisms, little is known regarding temperatures within coastal zones. Within coastal zones, seawater temperature varies as it is affected by a range of nearshore processes such as wave action, winds, tides, and turbidity. Additionally, the thermal properties of the substratum may also affect temperature variation (Smale and Wernberg 2009), allowing for a large range in variation at a localized scale. Given the significance of the impacts of temperature, it is imperative to understand how marine organisms may respond to temperature variation at coastal regions on both a global and local scale.

The coastal region of South Africa, spanning approximately 3113 km in length (Smit et al. 2013), has not yet been studied in great detail at highly localized scales. Upon looking on a broad scale, this region exhibits a large variation in seawater temperatures across its coastline (Mead et al. 2013; Smit et al. 2013) and is divided into three distinct coastal regions, each with contrasting temperatures. These include the cold temperate west coast, warm temperate south coast, and the subtropical east coast. These regions display noticeable differences in seawater temperatures in comparison to each other, primarily due to the influences of the neighbouring ocean currents (Bolton et al. 2004; Mead et al. 2013; Schlegel and Smit 2016). Existing temperature gradients are also associated with a difference in ecosystem physiology, species distribution and habitat structure (Smale and Wernberg, 2009 ; Wernberg et al 2011; Wernberg et al 2010). As a result of the diverse habitats along the coastline, species diversity is not uniformly distributed. The east and south coast has much higher diversity compared to the west coast (Mead et al 2013)

The South African coastline is bordered by the Benguela and Agulhas Currents on the west and east coasts respectively. The Benguela Current assists in transporting cold water northwards from the Southern Ocean (Hutchings et al. 2009; Schlegel et al. 2017), whereas the Agulhas, western boundary current, transports sub-tropical, warm water towards the tip of Africa (Schlegel et al. 2017). Together these two currents are responsible for the presence of a strong West-East thermal gradient occurring along the coastline of South Africa, with the west coast having significantly colder waters than the east coast. (Smit et al. 2013; Mead et al 2013). The south coast is unique as it affected by both the Benguela and Agulhas currents. In this way, the south coast region differs from the other two coastal regions as it experiences greater variations within its localized thermal gradient due to the influences of two contrasting ocean currents (Lutjeharms and Ballegooyen 1988). These currents can vary up to as much as 10°C (Schlegel et al 2017) within a 24-hour period thus creating fluctuations in seawater temperatures around the Western Cape region. The Agulhas Current is driven by a wind stress curl between the southeast trade winds and the Southern Hemisphere westerlies (Beal et al. 2011). The Agulhas current retroreflects after making contact with the Benguela current and may often interact with the South Atlantic (Hutchings et al., 2009), resulting in warm Indian Ocean water moving into the Atlantic at depths exceeding 2000m (Beal et al., 2011; Schlegel et al. 2017). These currents strongly affect the seawater temperatures around South Africa, and recent studies have shown that the temperatures have been changing. 

The statistical properties such as the mean, minimum, and maximum temperatures of *in situ* coastal seawater temperature time series representing the entire South African coastline, shows distinct coastal variations as a result of the Agulhas and Benguela currents (Schlegel and Smit 2016). Due to the small range of ambient temperatures experienced on the West coast, fluctuations are able to occur over a short period of time. Research has found that the seawater temperatures of the Benguela Current has been decreasing at a rate of approximately 0.5 °C per decade whilst the Agulhas Current has been increasing by between 0.55 °C - 0.7 °C per decade (Rouault et al. 2009; 2010). Overall, sea surface temperatures (SST) around South Africa have increased by approximately 0.25 °C between 1903 and 2013 (DEA, 2013) and are still increasing at a rate of 0.12 °C per decade (Schlegel and Smit, pers. comm.). The exact causes of these changes are not fully understood, however human activities are known to influence on coastal seawater temperature (Rouault et al. 2009, 2010 ; Mead et al. 2013, Mead 2011).

Due to their close proximity to populated or infrastructural areas, coastal regions are more accessible to humans in comparison to fully marine and open ocean regions. As a result, these areas are more susceptible to changes on both temporal and spatial scales(Help et al 2011; Kennish and Elliot 2011; Sink et al. 2012; Mead et al 2013). Along the South African coastline, direct pressures include exploitation of coastal resources, pollution and modification of communities (Griffiths et al 2004; Sink et al. 2012; Mead et al 2013, Mead 2011). These changes may directly be influenced by anthropogenic activity and hence promote climate change which ultimately influences air temperature, wind and sea surface temperature (Mead and Robinson 2011; Rouault and Mead 2011; Mead et al 2013).

Over the last few decades, improvements in remote sensing technology have enabled researchers to map global sea surface temperature with a high level of accuracy (Zainuddin et al. 2006; Smale and Wernberg 2009). The National Oceanic and Atmospheric Administration’s (NOAA) series of satellites have provided global SST datasets from the 1980s on both global and local scales (Pearce et al. 2005). The NOAA dataset is critically important as satellite temperature datasets are often used to monitor changes in oceanic temperatures and provide valuable information on both biological and physical parameters in the ocean (Demarcq et al. 2010). Furthermore, satellite-derived SST data plays an important role in the creating projections predicting the potential effects of climate change on coastal and oceanic marine biota (Muller et al. 2009; Wethey et al. 2011; Bartsch et al. 2012). Satellite-derived data is not as reliable as *in situ* temperature measurements, but are often used as a proxy when these measurements are scares or unavailable (Smale and Wernberg 2009). 

Satellite-derived temperature measurements often differ from *in situ* measurements (Katsaros 2003; Castillo and Lima 2010). This is primarily due to coastal satellite measurements failing to record small-scale variability or providing realistic information on average climatological temperatures. Smit et al. (2013) has shown that within offshore environments, satellite data generally reflects reality and has been validated against *in situ* measurements. Along coastal regions however, it was found that satellite-derived data had a warm bias of 6 °C when compared to coastal *in situ* collected seawater temperature data. For study sites located within a 400 m distance of the shoreline, it is recommended that data collected *in situ* be used over satellite-derived temperatures where possible (Smit et al. 2013).

Developments of *in situ* collected coastal temperature data products over the last three decades provide a more reliable source of accurate coastal seawater temperature data (Smit et al. 2013). The South African Coastal Temperature Network (SACTN) has collected SST data for the South African coastline from as early as 1972, with contributions from various organisations and governmental departments. This data was obtained via a combination of different instruments and methods, including hand held alcohol thermometers and underwater temperature recorders, and as a result of this inconsistency, the coastal seawater temperature data has been recorded at differing levels of precision (0.5°C and 0.001°C respectively) but are likely more reliable than satellite derived data.

It is important to understand the characteristics of the data sources which aid in yielding accurate and reliable measurements when comparing temperature variation along the coast. This led to the examination of the 129 *in situ* collected sites which makes up the SACTN dataset. The length of the time series and statistical characteristics of the dataset were used to guide a series of analysis. The length of the time series for each of the sites were adjusted systematically hence forming the core of the analysis. The variation of each of the time series differ as expected, by the Agulhas and Benguela Currents. The purpose of this study was to examine variations in temperature between selected sites along the South African coastline using seawater temperature data to better understand patterns of coastal temperature at a localized scale. Additionally, this study also examined the influence of environmental factors, namely, wave and wind action on coastal seawater temperatures. The aims of this study were to: i) examine whether there is homogeneity between the various sites sampled, ii) examine whether or not wind and wave action may cause a variation in seawater temperatures along the South African coastline and iii) to examine whether or not SST data collected via satellite may be affected by wind and wave action.

\newpage
# Methods

In order to compare abiotic parameters such as wind and wave action along the South African coastline,vast historical datasets for temperature, wave and wind were analysed and accessed. 

## Wave and wind data

Wind and wave action were important variables in this study as they are predicted to exhibit a direct influence on coastal water temperatures, and were used to investigate the impact Talarifyof abiotic parameters on seawater temperature at specific sites along the South African coastline. Wind and wave data was obtained from the South African Weather Service (SAWS), and were presented at three hour resolutions. The data was then used to model short–crested waves, generated by the wind into the coastal environment, using Simulating Waves in the Nearshore (SWAN). SWAN enables the extraction of wave parameters from specific gridded locations in the nearshore. A resolution of 200 meters was modelled, at both 7 and 15 meter contours.

## Temperature data

In order to measure variation in temperature across sites, the SACTN dataset was the primary source of temperature data. This dataset consisted of coastal seawater temperatures for 129 sites along the coast of South Africa, measured daily from 1972 until 2017. Of these, 80 were measured using hand-held thermometers and the remaining 45 were measured using underwater temperature recorders (UTRs).  Data collected predominately along the east coast of South Africa was obtained using glass thermometers from boats at the shark net locations along the coastline at a precision of 0.1 °C.

The SACTN dataset time series commenced from 1 January 1972 in Knysna, extending up to 20 December 2017 in Port Nolloth. Recordings per site were uneven, with the longest time series in the dataset being that of Gordons Bay, recorded by SAWS. Data collected for this region started on 13 September 1972 and concluded on 26 January 2017, with recordings still continuing daily. During the 1970s, a total of 11 time series began recording. A further 53 entries were added during the 1980s, 34 entries were added during the 1990s, and 18 entries were added during the 2000s. Recordings are still ongoing at several of these sites.

Seven sources contributed towards the compilation of this dataset: The Department of Environmental Affairs (DEA), South African Weather Services (SAWS), University of the Western Cape (UWC), Department of Forestry and Fisheries (DAFF), Kwa Zulu Natal Sharks Board (KZNSB), South African Environmental Observation Network (SAEON) and Ezemvelo KZN Wildlife Organization (EKZNW). KZNSB and SAWS made use of hand-held alcohol thermometers, taking measurements 20 cm below the sea surface. Temperature measurements taken with underwater temperature recorders, collected by DEA varied between a range of 0-28m, measurements obtained by DAFF varied between ranges of 0-9m, those collected by EKZN varied between ranges of 10-14m. SAEON and UWC placed the UTRs at depths of 0m and 0-6m respectively. Electronic instruments used for seawater temperature recordings have improved over the past four decades. Previous UTRs collected seawater temperature at a precision of 0.01°C whereas current models collect temperature at a thermal precision of 0.001°C. Data recorded using thermometers are saved and extracted annually.

## Data storage

Data collection protocols varied amongst the seven sources contributing to the SATCN dataset.  As a result, the data was not uniformly formatted. For these analyses, the data was combined and all temperature data were formatted into standardized comma delineated values (CSV) which allowed for a fixed methodology to be used across the entire dataset, thereby improving accuracy. Prior to data analysis, all data points exceeding 35 °C and/or below 0 °C were removed as these were  considered as outliers. These data points were then changed to NA (not available) so as to not interfere with analysis. All analyses performed in this thesis were conducted in R software version 3.4.2.

## Satellite derived SST data

This study made use of four satellite derived SST datasets in comparison with the SACTN *in situ* derived coastal seawater temperature dataset. The AVHRR- Only Optimum Interpolated Sea Surface Temperature (OISST) was used to determine SST within the study region. The AVHRR datasets have provided global SST’s for more than three decades (Reynolds and Smith, 1994; Pearce et al 2006). OISST is a global 1/4 ◦ gridded daily SST product that assimilates both remotely sensed and *in situ* sources of data to create a level-4 gap free product (Banzon et al., 201 6; Schlegel et al., 2017). The Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis, the second dataset, has been observed using satellite instruments with datasets spanning from 1st June 2002 to present times. MUR provides SST data at a spatial resolution of 0.01 degrees in longitude-latitude coordinates, at 1 km intervals; and is currently among the highest resolution SST available. The third dataset, a Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 sea surface temperature analysis, known as the K10 dataset, is produced at the Naval Oceanographic Office (NAVOCEANO) on a 0.1 degree grid, globally. This analysis makes use of SST observations from the AVHRR, the Geostationary Operational Environmental Satellite (GOES) Imager and the Advanced Microwave Scanning Radiometer for EOS (AMSR-E). The CMC dataset constitutes the forth dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 0.1 degree spatial resolution constructed by the Canadian Meteorological Center (CMC). This young dataset is available from the 1st January 2016. The CMC dataset combined infrared satellite SST at numerous points in the time series from the AVHRR, the European Meteorological Operational-A (METOP-A) and Operational-B (METOP-B) platforms, and microwave SST data from the Advanced Microwave Scanning Radiometer 2 in conjunction with *in situ* observations of SST from ships and buoys from the ICOADS program. For the statistical interpolation used to assimilate the satellite and *in situ* observations, the CMC incorporated the previous days’ analysis as the background field. The SST data used in this study were downloaded at a daily resolution within the latitude/longitude of the study region (Figure 1).

## Site selection

In order to compare temperatures along the South African coastline at a localized scale, appropriate sites from each of the major coastal regions needed to be selected. Since temperature data was not evenly recorded for each of the 129 sites representing South Africa’s coastline, the first step was to narrow down available sites that could be adequately compared. To do this a clustering analysis was perfomed using the `kmeans` function in R software with multiple random seeds to identify a number of clustering solutions that grouped sites together based on their available temperature data. The mean, minimum, and maximum temperature values were used within the clustering to group sites with similar temperatures along each coastal region. The clustering analysis represented the most accurate and distinct site groupings based on temperature distributions and yielded distinct east, south, and west coast groupings. Six clustering solutions along the South African coastline were also shown (Figure 1). With the data now divided into six distinct coastal regions, a time series was selected of at least one decade of data for each of the sites, excluding those with temperature data collected deeper than five meters. The length, decadal trend and precision of the time series were adjusted in a systematic manner to allow for meaningful comparisons to be made.
 
```{r , echo= FALSE, message=FALSE, warning=FALSE, fig.cap=" Figure 1: A map of the study area representing the 129 sites where *in situ* coastal seawater temperature was collected. These sites were grouped based on similar mean, minimum and maximum temperatures and as such each groups represents a unique colour variation.", fig.height=8, fig.width=12, fig.pos="H"}

knitr::include_graphics("figures/map_fixed.pdf")

```

Once sites were grouped together, the next step was to reduce the number of sites down to a manageable, but still representative, sub-sample of the whole. This was done primarily for two reasons. The first, simply, was to allow for the comparisons to be more readily interpretable by humans.  Secondly it was to allow for equal amount of sampling per coast. The east coast has previously been more heavily sampled than the rest and such an imbalance should be addressed. The criteria considered for the sub-samples included selecting the longest time series within the region and including data from as many different organizations as possible. This process yielded three sites for each of the clusters along the South African coastline (Figure 2). The statistical characteristics of the temperature were used to guide analysis of the time series to produce an accurate assessment of temperature variation between sites that were grouped together.
TalarifyTalarifyTalarifyTalarify
```{r run_fun2, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 2: A map representing the three sites chosen within each of the six cluster along the South African coastline.", fig.height=8, fig.width=12, fig.pos="H"}

knitr::include_graphics("figures/final_combined_plot.pdf")

```

With three sites per cluster, the haversine formula was used to calculate the geodesic distance between points specified by radians. Using this formula, the distances (km) between each of the sites along the coastline were determined. Thereafter, sites within the same cluster were matched based on the date that temperature was collected. This allowed for a comparison to test whether or not temperature variation exists between sites within the same cluster. Once the sites were matched, the means and standard deviations of temperatures between sites and clusters were determined. This highlighted the temperature variation between matched sites and allowed for seasonal comparisons within the same cluster.

Once the temperature variation between sites were carefully analysed, the seawater temperature data along with the wave and wind data were compared. Since the wave and wind data were collected at three hour resolutions, they were converted into daily data points in order to compare it with the temperature data. The wave data was collected at depths of 7 m or 15 m respectively. The ‘circular’ function in R software was then used to create circular objects around the wave data in order to calculate the daily wave and wind direction.

With temperature and wave values now corresponding to their respective sites, depths and dates, the hypothesis regarding whether or not a relationship existed between wind/wave action and temperature was tested. To do this, linear models for each site was produced, reflecting temperature and wave variations at each depth. Linear models typically produce coefficients of determination R2 as an output, the ‘purrr’ function within the ‘tidyverse’ R package was used to simultaneously compare temperature and wave data across sites and depths. Hereafter, ind and wave rose diagrams were constructed to determine the most predominant direction for a particular site. This was done to ascertain what the potential relationship between wind/waves and temperature was at each site during only the prevailing wind/wave directions.

## Satellite data analysis

Sea Surface temperature measurements used in this study were obtained from four different sources: MUR, CMC, K10 and AVHRR. A time series of SST was determined by creating a bounding box which represented the region of extent at the latitudes (-39.5° S, -25.5° S), and the longitudes (10.5° E, 39.5° E). The size of the pixel search area was set to use 2x the resolution of the sensor and at least 5x resol(km) from each of the stations. The satellite datasets and the corresponding SACTN *in situ* collected dataset were matched based on the latitudes, longitudes and the date at which temperature was collected. Some sites however, shared the same satellite data due to their close proximities. Once the satellite data corresponded with the *in situ* collected data, linear models for each site were produced, reflecting temperature and wave variations at each depth. Linear models typically produce coefficients of determination (R2 values) as an output, which is the statistic showing how much of the variance in a dependent variable is explained by the independent variable. This allows for us to test whether or not wave and wind direction may influence temperature at the various sites.

## Statistical analyses

A series of ANOVA tests were used to compare the main effects of the chosen variables on a continuous variable. In this analysis the relationship between matched sites based on the year and season as a function of the mean temperature were analysed, these analyses were used to test if significant differences occured between each pair of sites within each of the clusters. To compare whether temperature variation exist between sites found within the same clusters along the coast, boxplots were constructed. These plots enabled the visual identification of variations in temperature by summarising the descriptice statistics. To furthur analyse the temperature variation between sites line graphs were constucted. This plot allowed for visual identification of the variation in average temperature for each of the month and year for paired sites. All exploratory and statistical analyses were performed in RStudio version 1.1.442 (http://www.rstudio.com/). The data used within this study and comprehensive script used for data analyses, and production of figures are to be found at https://github.com/AmierohAbrahams/HONOURSPROJECT.

## Results

### Temperature variation

Temperatures were not uniformly distributed across the six clusters produced (Figure 3), with each set of sites having unique patterns of temperature variation. In cluster 1, which included Humburg, Eastern Beach and Orient Beach, it was found that temperature varied from approximately 13 ºC to 22 ºC. Within this cluster of sites, Hamburg had the highest maximum temperatures and the lowest minimum temperatures of the three sites. Conversely, Orient Beach had the lowest range of temperature variability, producing a comparatively short box plot. Orient Beach and Eastern Beach had relatively similar ranges and distributions of temperatures, as evident by their box plots nearly overlapping completely.

Within the cluster comprised of Mossel Bay, Stilbaai, and Knysna, temperatures ranged from approximately 12 ºC to 27 ºC, with most box plots being relatively long. Stilbaai had the widest range of temperature variation among the three sites but despite the apparent differences in temperature ranges between these sites, the average temperatures were relatively similar. Average temperatures were nearly completely identical within this cluster, with very few outliers present within the temperatures ranges of these sites.

Sites located within the third cluster had slightly lower temperatures than the previous two clusters. This cluster comprised of Bordjies, Saldanha, and Gansbaai and temperatures within here ranged from approximately 11 ºC to 21 ºC, with an average median temperature being close to 15 ºC across all three sites. Gansbaai had relatively low variation in temperature as it had a comparatively short box plot. Conversely, Saldanha had a long box plot representing high variation and relatively evenly distributed temperatures. These sites were similar in terms of their temperature variances, as their box plots were largely overlapping with few differences between them. There were however several outliers present within the temperatures of these sites.

The fourth cluster comprised of Port Edward, Leisure Bay, and T.O. Strand. Overall, the temperatures of these sites were higher than those of the sites within the other clusters, with a range of 15 ºC to 25 ºC, which is considerably higher.  The box plots for these sites were all decidedly long, representing a low variation of temperature with little skewness across sites. Temperatures did not differ between these three sites with each box plot overlapping very well. The median temperature for each of the sites within this cluster is 20.5 ºC.

Sites within the fifth cluster had overall lower temperatures than those within the remaining clusters. This cluster comprised of Port Nolloth, Lamberts Bay, and Sea Point, here sharp declines in average temperatures were observed throughout. Temperatures within this cluster ranges between 8 ºC and 18 ºC, with an average median temperature being close to 13 ºC. Port Nolloth had low variation in temperature as it had a comparatively short box plot with relatively evenly distributed temperatures. Lamberts Bay and Sea Point were similar in terms of temperature variances, as their box plots were largely overlapping with little differences between them. Several outliers were present within the temperatures of these sites.

In the cluster comprising of Kalk Bay, Muizenberg, and Gordons Bay the temperatures of these sites ranged from 8 ºC to 24 ºC, with most box plots being short. Muizenberg had the widest range of temperature variation of the three sites. Gordons Bay and Kalk Bay had identical temperature ranges. Similarly to the second cluster, despite the apparent differences in temperature ranges between these sites, the average temperatures across them were relatively similar and nearly identical.

```{r , echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 3: Boxplots representing the difference in seawater temperature of paired sites within each cluster along the South African coast.", fig.height=8, fig.width=12, fig.pos="H"}

knitr::include_graphics("figures/combined_plot.pdf")

```

On a monthly basis, the differences of average temperatures between sites within cluster 1, Eastern Beach, Orient Beach and Humburg,a large average temperature variation between Hamburg and Eastern Beach were seen during the summer and spring months of 1995 to 1997. For the remaining sites however, differences in average temperatures were low during Autumn. It was also evident that the average temperature between Hamburg and Orient Beach varied largely on an apparent seasonal basis. Slight average temperature variation existed between Eastern Beach and Orient Beach throughout the different seasons.

Converse to the first cluster, the cluster containing Mossel Bay, Knysna and Stilbaai, the largest differences in average temperatures were observed during autumn and winter months.  In this cluster large differences in average temperatures were present between Mossel Bay and Knysna, with the differences increasing annually from 1985 to 2017. Similarly, differences in average temperature also increased slightly between Stilbaai and Knysna during winter and spring. During summer months little differences in average temperature varation were seen between all three sites within this cluster.

Differences of average temperatures between sites within cluster 3 (Bordjies, Gansbaai and Saldanha) varied on an apparent seasonal basis. During the summer months we saw that there were large differences in average temperatures between Bordjies and the remaining three sites, with an increase in differences of average temperature between Saldanha Bay and Gansbaai throughout autumn, winter and spring.

In the fourth cluster, which comprised of Port Edward, Leisure Bay and T.O. Strand, small changes in the differences of average temperatures were notticed between sites on a monthly basis between 1980 to 2017. Here, the large differences in temperatures were observed towards the end of spring and during summer months. Temperatures were relatively stable throughout winter and the beginning of spring across all three sites found within this cluster.

In the cluster containing Lamberts Bay, Port Nolloth and Sea Point, large differences in average temperatures between sites at selected months between 1972 and 2017. During summer and Autumn months, differences in average temperature were observed between Lamberts Bay and the remaining sites increased. During the months of autumn Lamberts Bay and Seapoint showed large differences in average temperature variation. For the remaining sites, differences in average temperatures were relatively low throughout each month for the same time period.

In the cluster comprising of Kalk Bay, Gordons Bay and Muizenberg, the largest differences in average temperatures were observed during mid autumn and winter months. In this cluster large differences in average temperatures were seen between Muizenberg and the remaining sites, with the differences increasing annually throughout 1972 and 2016 during winter. Similarly, differences of average temperatures also increased between Kalk Bay and Muizenberg during these same months. In the summer and spring months little differences in average temperatures between sites, with minimal differences in the rates of these changes. These rates increased during spring.

```{r,  echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 4: Line graph representing the variation in average seawater temperature for each of the month and year for paired sites along the South African coastline.", fig.height=8, fig.width=12, fig.pos="H"}

knitr::include_graphics("figures/SACTN_clust_1_plot2.pdf")
knitr::include_graphics("figures/SACTN_clust_2_plot2.pdf")
knitr::include_graphics("figures/SACTN_clust_3_plot2.pdf")
knitr::include_graphics("figures/SACTN_clust_4_plot2.pdf")
knitr::include_graphics("figures/SACTN_clust_5_plot2.pdf")
knitr::include_graphics("figures/SACTN_clust_6_plot2.pdf")

```

### Average temperature of clustered sites

In the first cluster of sites, via the results of a one way ANOVA tests it was found that there was a significant differences in average temperatures between paired sites (F2 = 12.07, SS = 15.28, P < 0.001). These differences were present across season (F6 = 3.44, SS = 13.07, P < 0.002) but were not present yearly between individuals and paired sites (F2 = 1.38, SS = 1.75, P = 0.25). Similarly, paired sites within the second cluster also a significantly differed in average temperature (F2 = 166.84, SS = 418.6, P < 0.001). Conversely to first cluster however these differences were present yearly (F1 = 33.21, SS = 41.7, P < 0.001) and seasonally (F6 = 16.72, SS = 125.9, P < 0.02,) between individual and paired sites. In the third cluster of sites, ANOVA tests revealed that there were no significant differences in average temperatures between paired sites (F2 = 1.17, SS = 2.9, P = 0.31), but temperatures varied seasonally and yearly between individual sites (. In the fourth cluster there were again no significant difference of average temperatures between paired sites (F2 = 0.73, SS = 2.9, P = 0.48). Differences were also absent across seasons (F3 = 0.75, SS = 0.0042, P = 0.52) and years (F1 = 0.495, SS = 0.0009, P = 0.48) between both individual and paired sites. Sites within the fifth cluster were significantly different in average temperatures between paired sites (F2 = 77.10, SS = 196.7, P < 0.002), with these differences being present yearly (F1 = 172.80, SS = 220.4, P < 0.001) and seasonally (F3 = 77.10, SS = 29.92, P < 0.002) for both paired and individual sites. Finally, sites within the sixth cluster also significantly difference in average temperatures between paired sites, yearly and seasonally (F2 = 132.044, SS = 419.7, P <0.01).

## The coefficient of determination 

```{r, echo= FALSE, message=FALSE, warning=FALSE}
library(knitr)

SACTN_7_R2 <- load("~/Documents/HONOURSPROJECT/data/SACTN_15_R2.RData")
SACTN_15_R2 <- load("~/Documents/HONOURSPROJECT/data/SACTN_15_R2.RData")
AVHRR_temp_15_R2 <- load("~/Documents/HONOURSPROJECT/data/AVHRR_temp_15_R2.RData")
AVHRR_temp_7_R2 <- load("~/Documents/HONOURSPROJECT/data/AVHRR_temp_7_R2.RData")
MUR_7_R2 <- load("~/Documents/HONOURSPROJECT/data/MUR_7_R2.RData")
MUR_15_R2 <- load("~/Documents/HONOURSPROJECT/data/MUR_15_R2.RData")
CMC_7_R2 <- load("~/Documents/HONOURSPROJECT/data/CMC_7_R2.RData")
CMC_15_R2 <- load("~/Documents/HONOURSPROJECT/data/CMC_15_R2.RData")
K10_7_R2 <- load("~/Documents/HONOURSPROJECT/data/K10_7_R2.RData")
K10_15_R2 <- load("~/Documents/HONOURSPROJECT/data/K10_15_R2.RData")

kable(SACTN_7_R2, digits = 3, caption = "A table showing the impact of wind and wave action at a depth of 7 meters for the SACTN temperature dataset")
kable(SACTN_15_R2, digits = 3, caption = "A table showing the impact of wind and wave action at a depth of 15 meters for the SACTN temperature dataset")
knitr::kable(AVHRR_temp_7_R2, digits = 3, caption = "A table showing the impact of wind and wave action at a depth of 7 meters for the AVHRR temperature dataset")
knitr::kable(AVHRR_temp_15_R2, digits = 3, caption = "A table showing the impact of wind and wave action at a depth of 15 meters for the AVHRR temperature dataset")
knitr::kable(MUR_7_R2, digits = 3, caption = "A table showing the impact of wind and wave action at a depth of 7 meters for the MUR temperature dataset")
knitr::kable(MUR_15_R2, digits = 3, caption = "A table showing the impact of wind and wave action at a depth of 15 meters for the MUR temperature dataset")
knitr::kable(CMC_7_R2, digits = 3, caption = "A table showing the impact of wind and wave action at a depth of 7 meters for the CMC temperature dataset")
knitr::kable(CMC_15_R2, digits = 3, caption = "A table showing the impact of wind and wave action at a depth of 15 meters for the CMC temperature dataset")
knitr::kable(K10_7_R2, digits = 3, caption = "A table showing the impact of wind and wave action at a depth of 7 meters for the K10 temperature dataset")
knitr::kable(K10_15_R2, digits = 3, caption = "A table showing the impact of wind and wave action at a depth of 15 meters for the K10 temperature dataset")
```

# The impact of wind and wave action on temperature at a depth of 7meters

## SACTN:
The coeffcient of determination was calculated to determine the relationship between temperature and several variables including wind direction and speed as well as wave  height, period, direction and speed, for all 18 sites.  The results showed little variation in environmental factors and temperature for each of the sites. Wind and wave direction influenced temperature at 0 – 3%.  The most significant relationships were found in Muizenberg, Kalk Bay and Mossel Bay where the R^2 values indicated that wave period had a 6-7% influence on temperature.

## AVHRR:

Upon examining the impact of wind and wave action on seawater temperature, it was seen that wave and wind direction had a minimal effect on temperature variation with its R^2 values ranging between 0-3%. The results also indicated that wave height influenced temperature at some of the sites. This was evident in Gaansbaai and Lamberts Bay where wave height had a 9% impact on temperature variation.

## MUR:

The results continued to show little variation in regards to the influence of wind and wave action on temperature. At many of the sites, both wind and wave direction had a 0% impact on the temperature. However, it was seen that wave height and wave period had the greatest impact on temperature at some of the sites, with Gordons Bay and Gaansbaai indicating an 8% and 10% impact of wave height on temperature variation respectively.

## CMC:

The results indicated that wave as well as wind direction and wind speed showed the least significant impact on temperature variation with r^2 values ranging between 0-3%. Wave height continued to show the largest impact on temperature variation at some of the sites. Here, Gaansbaai and Gordons Bay indicated the highest R^2 values of 12% and 9% respectively.

## K10:

When comparing the impact of various environmental factors on temperature variation for the K10 data, the results indicated that the each of the above-mentioned variables had no impact on the temperature variation at Hamburg.  The impact of wind direction on temperature is highest at Gansbaai, representing an R^2 value of 4%. Wave height still repersented the greatest influence on temperature occurring at these sites

# The impact of wind and wave action on temperature at a depth of 15meters

## SACTN:

The coefficient of determination was calculated to determine the relationship between temperature and several variables including wind direction and speed as well as wave  height, period, direction and speed, for all 18 sites. The analyses indicated that there was no relationship present between the above-mentioned factors at Eastern Beach. Wave direction was seen to have no influence on temperature at any of the sites analysed as the R^2 values were at a constant 0%. Wind direction exhibited relationships between 0-4% across all sites, indicating that that factor did not significantly influence coastal temperature. The most significant relationships were found at the Muizenberg and Kalk Bay sites where the R^2 values indicated that wave period had an 8% and 7% influence on temperature respectively.  

## AVHRR:

The analyses indicated that a relationship existed between each of the environmental variables at Stilbaai with the intensity of these variables on temperature being extremely low. Wave period and wave height at Port Nolloth and Lamberts Bay represented the largest influence on temperature variation. This was still low with R^2 values between 6% - 8%.  In Gaansbaai wave height had a 9% influence on temperature, this being the largest value recorded.

## MUR:

Upon examining the relationship between the MUR SST dataset and each of the environmental factors it was found that wave height and wave period of each of the sites had the greatest influence on temperature variation compared to any of the other variables. Similarly to the AVHRR dataset,  Gaansbaai is largely influenced by wave height as indicated by R^2 value of 10%. Stillbaai and Seapoint also indicated that wave height had a 7% and 6% influence on temperature respectively. Muizenberg however, was the only site present where wave height had a 0% influence on temperature.

## CMC:

In the CMC SST dataset it was seen that wave height was the main factor influencing temperature at these sites. Gansbaai represented a R^2 value of 12% which was the highest value for each of the variables within this dataset. Each of the other environmental variables appeared to have a minor influence on the temperature for each of the sites ranging between 0% - 4%

## K10:

When comparing the relationship between SST of the K10 dataset and the environmental variables it was evident that Gaansbaai represented the largest R^2 value indicating a 7% influence of wave height on temperature. Wind direction exhibited the largest impact of variation, representing AN R^2 value ranging from 0 – 4%.

## Wind Rose diagram





## The impact of the most predominant wind and wave action on temperature variaiton



## Discussion

This study aimed to investigate how environmental factors such as wind and wave action influenced coastal seawater temperature variation along the South African coastline. Because seawater temperature is known to have large influences on species distributions (Bolton 2010; Smit et al. 2013), it is important to understand not only how temperatures vary across the coastline, but the factors driving these variations as well. Here, I confirmed that there were statistically significant differences in SSTs between 18 sites along the coastline of South Africa, with temperatures differing between coasts, and across sites along the same coasts. Overall, sites along the west coast had lower temperatures than those along the south and east coasts. I found that the east coast had overall little variation in temperatures between sites, whereas the south coast varied more frequently (Schlegel and Smit 2016).

My results confirmed that there were significant differences in temperatures between sites along the coastline, annually and seasonally, but did not provide evidence for which factors were driving these. Across each of the 18 sites that I assessed, I found that wind and wave action did not significantly influence SST. Variances in temperature between sites at different locations along the coastline were therefore not caused by variations in wind or wave action, suggesting that other factors were responsible for the observed patterns of temperature variation. These findings were consistent across six different datasets from various sources, with each of these providing no evidence of a relationship between wind or wave action and the SST of a site. However, these consisted of a combination of satellite and thermometer data, which varied in their measurements and time series.

Previous studies suggest that a longer time series containing more data would have greater accuracy in detecting subtle changes in temperature differences obtained from variable coastal regions such as the South African coastline. Schlegel and Smit (2016) suggested that having a time series of greater than 30 years for sites with a low variance results in increased ability to detect changes. They showed that high quality time series datasets have frequent measurements with minimal missing (NA) values present. Furthermore, low quality datasets (i.e. those with more than 5% of NA values) have a higher chance of detecting variation where none exists.

Datasets containing temperature information are inconsistent. Temperature collection along the South African coastline started in the 1970s, and has been inconsistent in terms of instruments used, modifications to styles and methods, calibration, and site locations (Smit et al. 2013; Schlegel and Smit 2016). Additionally, older records within some datasets, such as the SACTN dataset, may have been lost or are unreliable since the metadata for these are unavailable. These metadata are essential as their absence prevents us from fully understanding the influence that instruments had on temperature recordings. Some measurements may therefore not represent accurate temperatures due to the instrumentation used but we do however have an accurate indication of which recordings were measured with thermometers and which were with UTR.

The UTRs used to collect data in the SACTN dataset appeared to express a lower number of NA values compared to the data collected with hand thermometers. As such, this may have influenced the overall time series dataset (Smit and Schlegel 2016). The level of precision at which data was collected also influenced the length of the time series needed. Time series in which temperature were collected at a precision of 0.5 ºC may require another 24 months of recordings to precisely detect long term variation (Smit and Schlegel 2016). The average length of the thermometer time series component of the SACTN dataset was 346 months whereas the average length of UTR time series was less than half of that. With the extent of these differences in length being so severe, even once correcting for potential negative effects on the measurement precision of the thermometer collected time series, it was clear that thermometer data were more useful than that of UTRS.

Satellite acquired SST records are useful to modern marine scientists. This data is often used to model and predict a wide range of oceanic and biological processes in the open ocean but have only recently been used to study temperature variations influencing benthic organisms (Pearce et al. 2006). Here, both Satellite SST data, and in situ thermometer data were significantly correlated with wind and wave data along the coastal environment at a depth of seven and 15 meters respectively. SST data acquired by satellites are obtained from a thin boundary layer at the air-sea interface (Smale and Wernberg 2009) and thus deviated from coastal in situ collected seawater temperatures. These deviations may have affected the outcomes of my analyses thereby inhibiting by findings. However, when assessing the impacts of wind and wave action on satellite and in situ collected seawater temperature data I found the same trend of wind and wave action having insignificant impacts on the temperature variation along the South African coastline.

My findings were surprising but not unexpected. Along the coastline of South Africa, there is a known east-west thermal temperature gradient (Smit et al. 2013). This is caused by major oceanic processes such as coastal upwelling, thermohaline circulation, and the presence of major ocean currents, which cumulatively influences the temperatures along this coastline (Walker 1990; Smit and Schlegel 2016). While it was reasonable to assume that surface level environmental factors like wind and wave action would affect sea surface temperatures, it is not completely unexpected that they would have little effect given the prevalence of the major processes mentioned above. Those processes are undoubtedly the largest drivers behind coastal temperature variation and may simply be overpowering the effects of other environmental factors. 

Alternatively, it is possible that factors other than wind and wave action are influencing temperatures across the South African coastline. For example, we know that rainfall can have large influences on coastal SST (Reason et al. 1999) but other, non-climatic, factors could be playing a greater role. Coastal regions are highly impacted by human mediated pressures (Mead et al. 2013). These pressures are predicted to drive change over a spatial and temporal scale and is often a cause of temperature variation (Mead and Robinson 2011). Additionally, evidence indicates that human driven climate change on both a local and global scale largely influence seawater temperatures, wind regimes, and wave action (Rouault et al. 2009a, 2009b, 2010, 2011; Mead et al. 2013). These pressures are present along the South African coastline in the form of pollution, coastal runoff, invasive species, resource exploitation, and coastal mining, which directly and indirectly influences air temperature, wind, seawater temperature, and rainfall (Rouault and Mead 2011; Mead et al. 2013). These anthropogenic factors may be cumulatively playing an important role in affecting SST along the coastline, but the full extents of these are unknown. 

Within marine environments, coastal temperature variation allows for a variation in the spatial arrangements of marine biodiversity. Whilst wind and wave action may not be directly affecting ocean temperatures, Blamey and Branch (2008) have found that wave action has a profound influence on species distributions along the coastline. The presence or absence of marine species are determined by a variety of factors and whilst those factors may not be influencing each other as was the case here, they collectively play important roles in affecting the marine life of the South African coastline and identifying those roles can aid in improving our understanding of nearshore dynamics, thereby providing greater knowledge to be used for conservation. 

Here I have shown that wind and wave action are not directly affecting seawater temperature variation along the South African coastline, but that other factors may be. Future research could aim to examine the effects of air temperature and rainfall on the coastal seawater temperature for the 18 sites being assessed. Additionally, other factors such as the amount of sunlight penetrating the ocean or site exposure could also be tested, as well as assessing the optimal locations for data collection. Further studies should also consider examining how chlorophyll concentrations and salinity varies with temperature to assess the effects of seawater temperature on marine plant life. Homogenous coastlines with distinct temperature gradients such as in South Africa provide a model environment for temperature analyses at fine-resolutions. These data could provide critically important information that can be used to assist in conserving marine biodiversity within these waters and should be given greater focus within marine research in the near future.

## References

Bartsch, I., Wiencke, C., & Laepple, T. (2012). Global seaweed biogeography under a changing climate: the prospected effects of temperature. In Seaweed biology (pp. 383-406). Springer, Berlin, Heidelberg.

Beal, L. M., De Ruijter, W. P., Biastoch, A., Zahn, R., Cronin, M., Hermes, J., ... & Bornman, T. (2011). On the role of the Agulhas system in ocean circulation and climate. Nature, 472(7344), 429.

Bolton, J. J., Leliaert, F., De Clerck, O., Anderson, R. J., Stegenga, H., Engledow, H. E., & Coppejans, E. (2004). Where is the western limit of the tropical Indian Ocean seaweed flora? An analysis of intertidal seaweed biogeography on the east coast of South Africa. Marine biology, 144(1), 51-59.

Bolton, J. J. (2010). The biogeography of kelps (Laminariales, Phaeophyceae): a global analysis with new insights from recent advances in molecular phylogenetics. Helgoland Marine Research, 64(4), 263.

Griffiths, C. L., Van Sittert, L., Best, P. B., Brown, A. C., Clark, B. M., Cook, P. A., ... & Hutchings, K. (2005). Impacts of human activities on marine animal life in the Benguela: a historical overview. Oceanogr. Mar. Biol. Annu. Rev., 42, 303-392.

Griffiths, C. L., Mead, A., & Zietsman, L. (2011). Human activities as drivers of change on South African rocky shores. Observations on Environmental Change in South Africa, Sun Media, Stellenbosch, South Africa, 242-6.

Hutchings, L., Van der Lingen, C. D., Shannon, L. J., Crawford, R. J. M., Verheye, H. M. S., Bartholomae, C. H., ... & Fidel, Q. (2009). The Benguela Current: An ecosystem of four components. Progress in Oceanography, 83(1-4), 15-32.

Kennish M, Elliott M (eds). 2011. Vol. 8. Human-induced problems (uses and abuses). In: Wolanski E, McLusky D (chief eds), Treatise on estuarine and coastal science. Waltham: Academic Press.

Lutjeharms, J. R. E., & Van Ballegooyen, R. C. (1988). Anomalous upstream retroflection in the Agulhas Current. Science, 240(4860), 1770.

Mead, A. (2011). Climate and Bioinvasives drivers of change on South African Rocky shores? (Doctoral dissertation, University of Cape Town).

Mead, A., Griffiths, C. L., Branch, G. M., McQuaid, C. D., Blamey, L. K., Bolton, J. J., ... & Whitfield, A. K. (2013). Human-mediated drivers of change—impacts on coastal ecosystems and marine biota of South Africa. African Journal of Marine Science, 35(3), 403-425.

Mead A, Robinson T. 2011. Rocky shores and kelp beds. In: James N, Hermes J (eds), Insights into impacts of climate change on the South African marine and coastal environment. Pretoria: South African Environmental Observation Network (SAEON). Pp 69–76.

Müller, R., Laepple, T., Bartsch, I., & Wiencke, C. (2009). Impact of oceanic warming on the distribution of seaweeds in polar and cold-temperate waters. Botanica Marina, 52(6), 617-638.

Pearce, A., Faskel, F., & Hyndes, G. (2006). Nearshore sea temperature variability off Rottnest Island (Western Australia) derived from satellite data. International Journal of Remote Sensing, 27(12), 2503-2518.

Reason, C. J. C., & Mulenga, H. (1999). Relationships between South African rainfall and SST anomalies in the southwest Indian Ocean. International Journal of Climatology: A Journal of the Royal Meteorological Society, 19(15), 1651-1673.

Rouault M, Mead A. 2011. Trends in nearshore sea surface temperature and circulatory systems. In: James N, Hermes J (eds), Insights into impacts of climate change on the South African marine and coastal environment. Pretoria: South African Environmental Observation Network (SAEON). Pp 69–76.

Rouault, M., Penven, P., & Pohl, B. (2009). Warming in the Agulhas Current system since the 1980's. Geophysical Research Letters, 36(12).

Rouault, M. J., Mouche, A., Collard, F., Johannessen, J. A., & Chapron, B. (2010). Mapping the Agulhas Current from space: An assessment of ASAR surface current velocities. Journal of Geophysical Research: Oceans, 115(C10).

Schlegel, R. W., Oliver, E. C., Perkins-Kirkpatrick, S., Kruger, A., & Smit, A. J. (2017). Predominant atmospheric and oceanic patterns during coastal marine heatwaves. Frontiers in Marine Science, 4, 323.

Smale, D. A., & Wernberg, T. (2009). Satellite-derived SST data as a proxy for water temperature in nearshore benthic ecology. Marine Ecology Progress Series, 387, 27-37.

Smit, A. J., Roberts, M., Anderson, R. J., Dufois, F., Dudley, S. F., Bornman, T. G., ... & Bolton, J. J. (2013). A coastal seawater temperature dataset for biogeographical studies: large biases between in situ and remotely-sensed data sets around the coast of South Africa. PLoS One, 8(12), e81944. 

Sink K, Holness S, Harris L, Majiedt P, Atkinson L, Robinson T, Kirkman S, Hutchings L, Leslie R, Lamberth S, Kerwath S, von der Heyden S, Lombard A, Attwood C, Branch G, Fairweather
T, Taljaard S, Weerts S, Cowley P, Awad A, Halpern B, Grantham H, Wolf T. 2012. National Biodiversity Assessment 2011: Technical Report. Vol. 4: Marine and Coastal Component. Pretoria: South African National Biodiversity Institute.

Walker, N. D. (1990). Links between South African summer rainfall and temperature variability of the Agulhas and Benguela Current systems. Journal of Geophysical Research: Oceans, 95(C3), 3297-3319.

Wernberg, T., Thomsen, M. S., Tuya, F., Kendrick, G. A., Staehr, P. A., & Toohey, B. D. (2010). Decreasing resilience of kelp beds along a latitudinal temperature gradient: potential implications for a warmer future. Ecology letters, 13(6), 685-694.

Wernberg, T., Russell, B. D., Moore, P. J., Ling, S. D., Smale, D. A., Campbell, A., ... & Connell, S. D. (2011). Impacts of climate change in a global hotspot for temperate marine biodiversity and ocean warming. Journal of Experimental Marine Biology and Ecology, 400(1-2), 7-16.

Wethey, D. S., Woodin, S. A., Hilbish, T. J., Jones, S. J., Lima, F. P., & Brannock, P. M. (2011). Response of intertidal populations to climate: effects of extreme events versus long term change. Journal of Experimental Marine Biology and Ecology, 400(1-2), 132-144.

Zainuddin, M., Kiyofuji, H., Saitoh, K., & Saitoh, S. I. (2006). Using multi-sensor satellite remote sensing and catch data to detect ocean hot spots for albacore (Thunnus alalunga) in the northwestern North Pacific. Deep Sea Research Part II: Topical Studies in Oceanography, 53(3-4), 419-431.







